---
title: "Making & Indexing Data in R"
author: "Julia Parish"
date: "8/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(palmerpenguins)
library(tidyverse)

```

## Make vectors and check them out - have to be the same class (homogenous)

always think about what you want to have in code chunk vs console

```{r}
dogs <- c("teddy", 40, "waffle", "banjo", 5L)
dogs

# Check the class:
class(dogs)
```

# R will make assumptions on class of characters, but may not be the most useful. Check class of things and reassign as needed

```{r}
dog_age <- c(1.4, 5.2, 6, 10)
dog_age

class(dog_age)
typeof(dog_age)

dog_miles <- c(4L, 5L, 6L)
class(dog_miles)
typeof(dog_miles)

is.character(dog_miles)
is.character(dogs)

```
# Is.character can be built in as a unit test, to ensure you are using the correct vector. 

# Can assign names in vectors
```{r}
dog_food <- c(teddy = "purina", 
              khora = "alpo", 
              waffle = "fancy fest",
              banjo = "blue diamond")

dog_food
class(dog_food)

```
# To access different elements of a vector - reference & index

# To access info for khora
```{r}
dog_food[2]
dog_food["khora"]

# the outputs above are not stored, if you want to use them later, you must assign it

khora_food <- dog_food[2]
khora_food

```
# If you want to pull a range of elements

```{r}

cars <- c("red", "white", "yellow", "silver", "blue", "black", "purple", "green", "gray")

# reference a sequence of values
cars[2:5]

# you can have any variable name in the brackets

i <- 3

cars[i]

i <- c(1, 5, 7)
# i as 3 no longer exsists

cars[i]

# elements are mutable

cars[3] <- "BURRITOS"

```

## Lists in R

```{r}
urchins <- list("blue", 2, FALSE, 1200L)
urchins 
# tells you different elements in list when you call it [[1]], [[2]], etc...

# actual value - contents of the element; almost always want to use this [[]] so you are calling the actual content
urchins[[1]]

# actual class - contents is list
urchins[1]

# 
urchins <- list("blue", 
                c(1 ,2, 3), 
                seq(from = 0, to = 50, by = 0.5), 
                1200L)
urchins
```

# name things in list - each element has the same number of things making it pretty close to a data frame
```{r}
tacos <- list(topping = c("onion", "cabbage", "avocado"),
              filling = c("beans", "fish", "meat"),
              price = c(6.75, 8.25, 9.50)) 
tacos

# to get taco prices - not a list
taco_price <- tacos[[3]]
taco_price

# if you have a dataframe, how do you call to a specific column? Use $ OR [[]]

tacos$filling

tacos[["filling"]]
              
```

## Attach palmer penguins and tidyverse! All work above has been in base R. 
### Now to make a data frames - should feel like making a list

```{r}
fruit <- data.frame(type = c("apple", "banana", "peach"),
                    mass = c(130, 195, 150))
fruit
class(fruit)

# to refer to a single piece of data frame, use $; not typically storing in Rmd files, but should always be checking, esp if math isn't working
class(fruit$mass)

# to store value of banana mass - the value is 2,2 (2nd row, 2nd column) To get the entire 'type' column back - use the $, fruit$type OR fruit[, 1] *don't forget the comma! It is saying get all rows in column 1. Typically do not want to hard code a position

fruit[, 1]
fruit$type

```







