---
title: "penguin_filtering"
author: "Julia Parish"
date: "8/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
```

# I want to create a scatterplot of flipper length versus body mass, with body mass converted to kilograms, only for Adelie penguins. - need to add a column to convert grams to kilograms so divide by 1000. grams(1 kg/1000g) always write down conversions by hand

creating a new object stored in enviro as "adelies" (and then = pipe command - use shortcut, then press enter to vertically arrange) Using Filter - don't have to add data name b/c you piped into the filter function, give it a conditional statement to only keep adelie penguins, to find out if it is a match use == (it is case senstive), filter is similar to subset (base R) but plays nicer with tidyverse functions
# to add a new column, use the mutate function, you are assigning it so =, not asking a ? when we used ==

what does facet wrap do for ggplot? Creates new panels in plot based on a factor or element (ex: island)
```{r}
adelies <- penguins %>% 
  filter(species == "Adelie") %>% 
  mutate(body_mass_kg = body_mass_g / 1000)

ggplot(data = adelies, aes(x = flipper_length_mm, y = body_mass_kg)) + 
  geom_point(aes(color = island)) + 
  facet_wrap(~island)

unique(adelies$species)
```

# no facet wrap of above ggplot
```{r}
ggplot(data = adelies, aes(x = flipper_length_mm, y = body_mass_kg)) + 
  geom_point(aes(color = island))

# answered Desik's Q re: aes
ggplot(data = adelies, aes(x = flipper_length_mm, y = body_mass_kg)) + 
    geom_point(aes(color = "blue")) +
    geom_line(aes(color = island)) + 
    facet_wrap(~island)

ggplot(data = penguins) +
   geom_point(aes(x = flipper_length_mm, y = body_mass_g))


```

